RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-11-09 introduction.tex
Update introduction.tex
very small change (mention semantics rights after reqs)

Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}). Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason. Associated with argumentation schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way. We believe that argument schemes are are a good choice for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.

@@

In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models. In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}. We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.

>>>>

Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}. Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason. Associated with argument schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way. These critical questions can point to counterarguments, and the acceptability of sets of arguments can be computed using argumentation semantics. Thus computational argumentation based on schemes can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.

@@

In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models. In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}. We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
